---
title: "Does Location influence startup success" 
output:
  pdf_document: default
  html_document: default
---

## Reading the CSV


```{r}
df <- read.csv("C:/Users/chaki/Downloads/cleaned_investments_VC.csv")
```

```{r}
head(df)
```

```{r}
nrow(df)
```

# Country level analysis

## Data Manipulation

```{r}
library(tidyverse)
top_5_countries <- df %>%
  count(country, sort = TRUE) %>%
  head(5) %>%
  pull(country)

df$country[!(df$country %in% top_5_countries)] <- "Others"
head(df)
```

## Visualising success status proportion (as percentage) for countries

```{r}
library(ggplot2)

ggplot(df, aes(x = country, fill = success_metric_updated)) +
  geom_bar(stat = "count", position = "fill") +
  scale_fill_brewer(palette = "Set3") +
  scale_y_continuous(name = "percent",
                     breaks = c(0, 0.5, 1),
                     labels = scales::percent(c(0, 0.5, 1))) +
  labs(title = "Proportion of Success Metrics by Country",
       x = "Country",
       y = "Proportion") +
  # Increase font size and rotate x-axis labels
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        axis.text.x = element_text(angle = 45, hjust = 1))
```

## Merging "Potentially Successfully" into "Unsuccessful"

```{r}
df_new <- df
df_new$success_metric_updated[df$success_metric_updated == 'Potentially Successful'] <- 'Unsuccessful'
head(df)
```

```{r}
df_new <- df_new[c("country", "success_metric_updated")]
head(df_new)
```

## Contigency tables

```{r}
tbl <- table(df_new$country,df_new$success_metric_updated)
tbl
```


```{r}
library(lsr)


cramer_v <- cramersV(tbl)

cat("Cramer's V:", cramer_v)
```

```{r}
tbl_prop <- prop.table(tbl, 1)
tbl_prop
```

## Chi square

```{r}
chisq.test(tbl)
```

# Testing if proportion of success for US is different from other countries

```{r}
# Create a new data frame with the "Others" category
df_new_merge <- df[c("country", "success_metric_updated")]
df_new_merge$country[!(df_new_merge$country %in% c("United States", "Others"))] <- "Others"

# Create a new table with the updated data frame
tbl_merge <- table(df_new_merge$country, df_new_merge$success_metric_updated)

tbl_merge
```

```{r}
cramer_v <- cramersV(tbl_merge)
cat("Cramer's V:", cramer_v)
```

```{r}
# Subset data for US and Others
usa_data <- df_new_merge[df_new_merge$country == "United States", ]
other_data <- df_new_merge[df_new_merge$country != "United States", ]

# Calculate success proportions for each group
p_usa <- sum(usa_data$success_metric_updated == "Successful") / nrow(usa_data)
p_other <- sum(other_data$success_metric_updated == "Successful") / nrow(other_data)

# Calculate pooled proportion (assuming equal variances)
pooled_p <- (sum(usa_data$success_metric_updated == "Successful") + sum(other_data$success_metric_updated == "Successful")) / (nrow(usa_data) + nrow(other_data))

# Calculate standard error for the difference in proportions
se_diff <- sqrt(pooled_p * (1 - pooled_p) * (1/nrow(usa_data) + 1/nrow(other_data)))

# Calculate the z-statistic
z_stat <- (p_usa - p_other) / se_diff

# Interpretation
# A p-value less than the chosen significance level 0.05 indicates
# a statistically significant difference between the success proportions
# in US and Others category.
```

```{r}

# Confidence level (e.g., 95%)
confidence_level <- 0.95

# Critical value (one-tailed) from standard normal distribution table for confidence level
z_crit <- qnorm(1 - (1 - confidence_level) / 2, mean = 0, sd = 1)

# Confidence interval for the difference in proportions
lower_bound <- (p_usa - p_other) - z_crit * se_diff
upper_bound <- (p_usa - p_other) + z_crit * se_diff

# Interpretation
# Report the z-statistic, p-value, and confidence interval.
# The confidence interval captures the range of plausible values for the true
# difference in success proportions between US and Others with a confidence level
# of (e.g.,) 95%.

cat("z-statistic:", z_stat, "\n")
cat("p-value (two-tailed):", 2 * pnorm(-abs(z_stat), mean = 0, sd = 1, lower.tail = TRUE), "\n")
cat("Confidence Interval (", confidence_level * 100, "%):", lower_bound, " - ", upper_bound, "\n")
```

# City level analysis

```{r}
df_us <- df %>%
  filter(country == "United States")

length(unique(df_us$city))

top_5_cities <- df_us %>%
  count(city, sort = TRUE) %>%
  head(5) %>%
  pull(city)

df_us$city[!(df_us$city %in% top_5_cities)] <- "Others"
head(df_us)
```

```{r}
library(ggplot2)

ggplot(df_us, aes(x = city, fill = success_metric_updated)) +
  geom_bar(stat = "count", position = "fill") +
  scale_fill_brewer(palette = "Set3") +
  scale_y_continuous(name = "percent",
                     breaks = c(0, 0.5, 1),
                     labels = scales::percent(c(0, 0.5, 1))) +
  labs(title = "Proportion of Success Metrics by City",
       x = "City",
       y = "Proportion") +
  # Increase font size and rotate x-axis labels
  theme(axis.text = element_text(size = 14),  # Change 14 to your desired size
        axis.title = element_text(size = 16),  # Change 16 to your desired size
        axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

```

```{r}
df_us_new <- df_us
df_us_new$success_metric_updated[df_us$success_metric_updated == 'Potentially Successful'] <- 'Unsuccessful'
```

```{r}
df_us_new <- df_us_new[c("city", "success_metric_updated")]
head(df_us_new)
```

```{r}
tbl2 <- table(df_us_new$city,df_us_new$success_metric_updated)
tbl2
```

```{r}
cramer_v <- cramersV(tbl2)

cat("Cramer's V:", cramer_v)
```

```{r}
tbl_prop2 <- prop.table(tbl2, 1)
tbl_prop2
```

```{r}
chisq.test(tbl2)
```

```{r}
# Create a new data frame with the "Others" category
df_us_new_merge <- df[c("city", "success_metric_updated")]
df_us_new_merge$city[!(df_us_new_merge$city %in% c("Seattle", "Others"))] <- "Others"

# Create a new table with the updated data frame
tbl_us_merge <- table(df_us_new_merge$city, df_us_new_merge$success_metric_updated)

tbl_us_merge
```

```{r}
# Subset data for US and Others
seattle_data <- df_us_new_merge[df_us_new_merge$city == "Seattle", ]
other_data <- df_us_new_merge[df_us_new_merge$city != "Seattle", ]

# Calculate success proportions for each group
p_sea <- sum(seattle_data$success_metric_updated == "Successful") / nrow(seattle_data)
p_other <- sum(other_data$success_metric_updated == "Successful") / nrow(other_data)

# Calculate pooled proportion (assuming equal variances)
pooled_p <- (sum(seattle_data$success_metric_updated == "Successful") + sum(other_data$success_metric_updated == "Successful")) / (nrow(seattle_data) + nrow(other_data))

# Calculate standard error for the difference in proportions
se_diff <- sqrt(pooled_p * (1 - pooled_p) * (1/nrow(seattle_data) + 1/nrow(other_data)))

# Calculate the z-statistic
z_stat <- (p_sea - p_other) / se_diff

# Interpretation
# A p-value less than the chosen significance level 0.05 indicates
# a statistically significant difference between the success proportions
# in US and Others category.
```

```{r}

# Confidence level (e.g., 95%)
confidence_level <- 0.95

# Critical value (one-tailed) from standard normal distribution table for confidence level
z_crit <- qnorm(1 - (1 - confidence_level) / 2, mean = 0, sd = 1)

# Confidence interval for the difference in proportions
lower_bound <- (p_sea - p_other) - z_crit * se_diff
upper_bound <- (p_sea - p_other) + z_crit * se_diff

# Interpretation
# Report the z-statistic, p-value, and confidence interval.
# The confidence interval captures the range of plausible values for the true
# difference in success proportions between US and Others with a confidence level
# of (e.g.,) 95%.

cat("z-statistic:", z_stat, "\n")
cat("p-value (two-tailed):", 2 * pnorm(-abs(z_stat), mean = 0, sd = 1, lower.tail = TRUE), "\n")
cat("Confidence Interval (", confidence_level * 100, "%):", lower_bound, " - ", upper_bound, "\n")
```

# Additional analysis: Logistic Regression

## Country level analysis

```{r}
df_new$success_metric_updated <- as.numeric(df_new$success_metric_updated == "Successful")
```

```{r}
df_new <- within(df_new,relevel(factor(country),ref = "Others"))
head(df_new)
```

```{r}
model <- glm(success_metric_updated ~ country , data = df_new, family = "binomial")
summary(model)
```

```{r}
library("effects")
plot(Effect('country', model), confint=T)
```

```{r}
# Get the estimated coefficients
coefs <- coef(model, simplify=TRUE)

# Calculate odds ratios for each country compared to the reference category
odds_ratios <- as.data.frame(exp(coefs))
odds_ratios
```

## City level analysis

```{r}
df_us_new <- within(df_us_new,relevel(factor(city),ref = "Others"))
```

```{r}
library(nnet)
df_us_new$city <- as.factor(df_us_new$city)
df_relevel_us <- df_us_new %>%
  mutate(city = relevel(city, ref = "Seattle"))
model2 <- multinom(success_metric_updated ~ city, data = df_relevel_us)
summary(model2)
```

```{r}
df_us_new$success_metric_updated <- as.numeric(df_us_new$success_metric_updated == "Successful")
```

```{r}
model2 <- glm(success_metric_updated ~ city, data = df_us_new, family = "binomial")
summary(model2)
```

```{r}
# Get the estimated coefficients
coefs <- coef(model2, simplify=TRUE)

# Calculate odds ratios for each country compared to the reference category
odds_ratios <- as.data.frame(exp(coefs))
odds_ratios
```

```{r}
library("effects")
plot(Effect('city', model2), confint=T)
```

